---
title: "Molecular Heatmap - Exact Feature Match"
author: "Jordan Lee"
date: "2/2/2021"
output: html_document
---

# TMP GDAN working group 
## Figure 4 Suggestion

### Question

We see (in other figures) that different teams have similar high classification performance. Is this similar performance due to selection of the same feature sets?

### Rationale

If we see teams picking different features, this suggests that there are multiple biomarkers that can predict cancer subtype well. An alternative explanation if we see teams picking different features, is that perhaps there is some relationship (correlation, biological pathway connection, etc.) between features so in a sense the teams are picking the "same" features if they are correlated.

In this analysis we will address: are groups picking the same features? What is the relationship?


### Analysis


The preprocessing and generation of these files was conducted in a custom python pipeline. A summary of the pipeline is:

1. Pick a cancer cohort (here we picked breast cancer)
2. Select the best model from each of the 5 teams (JadBIO, SubSCOPE, AKLIMATE, Sci-Kit Grid, and Cloud Forest). Based on highest mean overall weighted F1 score
3. Pull the corresponding feature set from each model
4. Map features to molecular file (based on Theo's tarball)
5. Cluster feature columns, no clustering of sample rows Clustering done on molecular file

Scaling data - METH, GEXP, and MIR

Non-scaled data - MUTA, CNVR

I want to see the differences between features. Data was scaled - all samples for a given feature have mean 0 and sd 1. then ran through normal pipeline

Note in calculating z-scores, NAs were omitted

```{r}
###### 
# Hardcoded 
######
cancer <- 'BRCA' # 'GEA', 'LGGGBM', 'BRCA'

outdir ='../data/figure_panel_b/'

file_imp_aklimate <- paste(
  '../src/aklimate_feature_importance_scores_20200807/AKLIMATE_TEMPLATE_FULLY_TRAINED_', 
  cancer, 
  '_20200423_aklimate_ranked_feature_importance.tsv', 
  sep=''
)

jd_file <- paste(
  '../src/feature_importance/JADBio_TCGA_results_fig6_update_2020_01_allcohorts_MCastro/', 
  cancer, 
  '_markers_outcome_association.csv', 
  sep = ''
)

yes_scale <- c('N:METH', 'N:GEXP','N:MIR') # which fts to scale

#####
# Packages and Functions
#####
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(testit))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(docstring))


model2team <- function(df){
  #' Read team model names and map to team name.
  #' Returns named vector of specific team models.
  #' Names: "JADBIO", "CForest", "AKLIMATE", "SubSCOPE", "SKGrid"
  headers_vector <- c()
  names_vector <- c()
  models <- colnames(df)[!colnames(df) %in% c('featureID', 'Total')]
  
  for (header in models){
    if (grepl('gnosis', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'JADBIO')
    } 
    else if (grepl('CF', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'CForest')
    }
    else if (grepl('nn_jg', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'SubSCOPE')
    }
    else if (grepl('AKLIMATE', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'AKLIMATE')
    }
    else {
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'SKGrid')
    }  
  }
  names(headers_vector) <- names_vector
  return(headers_vector)
}


get_colors <- function(df){
  #' Pull color codes based on total number of subtypes found in "Labels" column
  #' Input df that contains column "Labels" where subtypes are
  #' Returns vector of set heatmap color codes for subtypes
  nsubs <- df %>% 
    select(Labels) %>% 
    unique() %>%
    nrow() %>% 
    as.integer()
  if (nsubs == 4){
    color_codes <- c(
      "1" = 'salmon4',
      "2"='red3',
      "3"='orangered',
      "4"='orange1'
    )
  } else if (nsubs == 7){
    color_codes <- c(
      "1" = '#F58748',
      "2" ='#A95757',
      "3"='#4A2918',
      "4"='red3',
      "5"='orangered',
      "6"= 'salmon',
      "7" = 'orange1'
    )  
  }
  return(color_codes)
}



get_base_heatmap <- function(prefix, cancer, header_jadbio, header_cforest, header_aklimate, header_subscope, header_skgrid){
  #' Create base heatmap - no hallmark info. For exploratory purposes
  ######
  # Preprocess
  #####
  # A. Order by subtype
  df_transform <- df %>% arrange(Labels)
  # B. Column annotation
  subtypestring <- paste(cancer, '_', sep='')
  subtype_ha <- rowAnnotation(
    Subtype = 
      select(df_transform, Labels) %>%
      mutate_all(funs(str_replace(., subtypestring, ""))) %>%
      as.matrix(), 
    na_col = 'grey', 
    col = list(
      Subtype = get_colors(df)
    ),
    show_annotation_name = FALSE,
    simple_anno_size = unit(3, "mm") # width
  )
  # C. Select data type
  df_transform <- df_transform %>%
      select(-Labels) %>%
      select(-all_of(cancer)) %>%
      select(starts_with(prefix))
  mat <- df_transform %>%
      as.matrix() %>%
      t()
  print(prefix)
  n_fts <- nrow(mat)
  print(n_fts)
  
  #####
  # 1. Generate temp Heatmap 
  # and pull row/col order
  #####
  # Scale if appropriate 
  #z-scores == each ft row will have mean 0, sd 1. omit NAs
  if (prefix %in% yes_scale){
    mat <- scale(t(mat), center=TRUE, scale=TRUE) 
  } else {
    mat <- t(mat) #flip for heatmap looks
  }
  # Heatmap
  fig <- Heatmap(
    mat,
    name = 'first heatmap',
    cluster_rows = FALSE,
    clustering_distance_columns = "euclidean",
    clustering_method_columns = "ward.D",
    cluster_columns = TRUE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = paste('Features (n=', n_fts, ')', sep=''),
    row_title = paste('Samples (n=', nrow(mat), ')', sep=''),
    right_annotation = subtype_ha
  )
  #### 
  # Add team annotation bar
  ####
  # Ordering
  # 1. Get order of features post heatmap clustering
  heatmap_order <- column_order(fig) # index vector
  ftnames_order <- c() # featurename vector
  for (i in heatmap_order){
    add_ft <- colnames(mat)[i]
    ftnames_order <- c(ftnames_order, add_ft)
  }
  # 2. Get new matrix that is ordered by heatmap clustering
  mat2 <- mat[,match(ftnames_order, colnames(mat))]
  #####
  # Build annotation bars of teams feature sets
  #####
  # 1. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% 
    filter(featureID %in% ftnames_order) %>% 
    arrange(match(featureID, ftnames_order))
  # 2. Pull just the team of interest
  jadbio <- team_df %>% 
    pull(header_jadbio) %>% 
    as.character()
  cforest <- team_df %>% 
    pull(header_cforest) %>% 
    as.character()
  aklimate <- team_df %>% 
    pull(header_aklimate) %>% 
    as.character()
  subscope <- team_df %>% 
    pull(header_subscope) %>% 
    as.character()
  skgrid <- team_df %>% 
    pull(header_skgrid) %>% 
    as.character()
  team_list <- HeatmapAnnotation(
    JADBio = jadbio,
    CloudForest = cforest,
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    SciKitGrid = skgrid,
    col = list(
      JADBio = c('0' = "#333333", '1' = "#D55B5B"),
      CloudForest = c('0' = "#333333", '1' = "mediumpurple1"),
      AKLIMATE = c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE = c('0' = "#333333", '1' = "palegreen2"),
      SciKitGrid = c('0' = "#333333", '1' = "#EFA9A9")
    ),
    show_legend = FALSE,
    nTeamsSelected= anno_barplot (
      team_df$Total,
      bar_width=1
    )
    # simple_anno_size = unit(2, 'mm') # height
  )
  #####
  # 3. Output Heatmap
  #####
  # Set up saving fig packet
  tiff(
    paste(
      cancer,
      '_ht_base_', 
      unlist(strsplit(prefix, ':'))[2],
      '.tiff',
      sep=''
    ),
    width = 1400,
    height = 1200,
    res = 200,
    compression = "none"
  )
  # Draw
  fig <- Heatmap(
    mat2,
    name = prefix,
    # width = unit(12, 'cm'),
    # height = unit(12, 'cm'),
    cluster_rows = FALSE,
    clustering_distance_columns = "euclidean",
    clustering_method_columns = "ward.D",
    column_order = ftnames_order,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = paste('Features (n=', n_fts, ')', sep=''),
    row_title = paste('Samples (n=', nrow(mat2), ')', sep=''),
    right_annotation = subtype_ha,
    bottom_annotation = team_list,
    na_col = 'white'
  )
  draw(fig)
  dev.off()
  #####
  # unit tests
  #####
  # sanity check 1 - 
  #that the row orders are the same in the two heatmaps
  assert('Assertion Error: order of feature rows on 
         heatmap and final heatmap not same', 
         colnames(mat2) == ftnames_order
  )
  # sanity check 2 - 
  #directly compare heatmap rows
  heatmap_order <- column_order(fig) # index vector
  ftnames_order2 <- c() # featurename vector
  for (i in heatmap_order){
    add_ft <- colnames(mat2)[i]
    ftnames_order2 <- c(ftnames_order2, add_ft)
  }
  assert('Assertion Error: order of feature rows on 
         heatmap and final heatmap not same', 
         ftnames_order2 == ftnames_order
  )
  print(
    paste(
      'Distance metric = ', 
      fig@column_dend_param$distance, 
      '. Method = ', 
      fig@column_dend_param$method, 
      sep=' '
    )
  )
  # Assign to 'output' variables
  return(
    list(
      'results_matrix' = mat2, 
      'results_ft_order' = ftnames_order, 
      'subtype_annotation' = subtype_ha
    )
  )
}


build_hallmark_vect <- function(hallmark, ftnames_order, platform_of_interest){
  #' Create hallmark vector
  #' for builidng hallmark heatmap
  fts_checked <- c() # sanity check
  hallmark_present <- c()
  for (feature in ftnames_order){
    # 1. Preprocess - to gene symbol
    if (platform_of_interest == 'N:GEXP'){
      GENE <- unlist(strsplit(feature, '::'))[2]
      GENE <- unlist(strsplit(GENE, ':'))[1]
    } else if (platform_of_interest == 'N:METH'){
      GENE <- unlist(strsplit(feature, ':'))[4]
    }
    # 2. Hallmark Mapping
    halls <- mappings[mappings$human_gene_symbol==GENE,]$gs_name %>% 
      as.vector()
    # If hallmark present
    if (hallmark %in% halls){
      i <- match(hallmark, halls)
      hallmark_present <- c(hallmark_present, 1)
      fts_checked <- c(fts_checked, feature)
    }
    else {
      hallmark_present <- c(hallmark_present, 0)
      fts_checked <- c(fts_checked, feature)
    }
  }
  return(hallmark_present)
}


mm_normal <- function(x) {
  #' Calculate min-max normalization of input values
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
####
# Read in files
####
# A. base files
df <- fread(
  paste(
    '../data/figure_panel_b/', 
    cancer, 
    '_fts_by_VALUE.tsv', 
    sep=''
  )
) %>% as.data.frame()

df_fts <- fread(
  paste(
    '../data/figure_panel_b/', 
    cancer, 
    '_fts_by_TEAM.tsv', 
    sep=''
  )
)%>% as.data.frame()

mappings <- fread(
  '../data/figure_panel_b/hallmarks.tsv'
) %>% as.data.frame()

# B. Importance Score files
imp_aklimate <- fread(
  file_imp_aklimate
) %>% as.data.frame()

imp_jd <- fread(
  jd_file
) %>% as.data.frame()
colnames(imp_jd) <- c("features","importance")

####
# Define platform for hallmark heatmap
####
if (cancer == 'BRCA'){
  platform_of_interest <- 'N:GEXP'
} else if (cancer == 'LGGGBM'){
  platform_of_interest <- 'N:METH'
} else if (cancer == 'GEA'){
  platform_of_interest <- 'N:METH'
}

####
# Build list of data types present
####
if (cancer == 'BRCA' | cancer == 'GEA' ){
  platforms <- c('N:METH', 'N:MIR', 'I:CNVR', 'B:MUTA', 'N:GEXP')
} else if (cancer == 'LGGGBM'){
  platforms <- c('N:METH', 'I:CNVR', 'B:MUTA', 'N:GEXP')
}
```

### Section 1: Heatmaps for each data type

Code and figures for all data types

```{r mixed scaled z score loop for all data types}
setwd(outdir)

####
# Create Base Heatmap for ALL DATA TYPES
####
# Get models
models <- model2team(df_fts)
# Set up saving fig packet
for (prefix in platforms){
  # A. Base Heatmap
  results_list <- get_base_heatmap(
    prefix, 
    cancer, 
    models['JADBIO'],
    models['CForest'],
    models['AKLIMATE'],
    models['SubSCOPE'],
    models['SKGrid']
  )
  # B. If platform for Hallmarks Heatmap then save global variables - input for Section 2
  if (prefix == platform_of_interest){
    print('Using this as main platform of interest:')
    print(prefix)
    mat2 <- results_list[['results_matrix']]
    ftnames_order <- results_list[['results_ft_order']]  
    subtype_annotation <- results_list[['subtype_annotation']]
    n_fts <- length(ftnames_order)
  }
}
```

### Section 2: Explore the hallmarks associated with GEXP features

One annotation bar (top hallmarks only, all others are not colored)

```{r explore hallmarks}
setwd(outdir)

######
# Explore Hallmarks
######
# 1. Prep
#look up if in hallmark mappings table
n_hallmarks <- c() 
pooled_hallmarks <- c() 

if (platform_of_interest == 'N:GEXP'){
  for (feature in ftnames_order){
    # 1. Preprocess - to gene symbol
    GENE <- unlist(strsplit(feature, '::'))[2]
    GENE <- unlist(strsplit(GENE, ':'))[1]
    # 2. Hallmark Mapping
    halls <- mappings[mappings$human_gene_symbol==GENE,]$gs_name %>% as.vector()
    # Multiple Hallmarks
    n_hallmarks <- c(n_hallmarks, length(halls))
    pooled_hallmarks <- c(pooled_hallmarks, halls)
  }  
} else if (platform_of_interest == 'N:METH'){
  for (feature in ftnames_order){
    # 1. Preprocess - to gene symbol
    GENE <- unlist(strsplit(feature, ':'))[4]
    # 2. Hallmark Mapping
    halls <- mappings[mappings$human_gene_symbol==GENE,]$gs_name %>% as.vector()
    # Multiple Hallmarks
    n_hallmarks <- c(n_hallmarks, length(halls))
    pooled_hallmarks <- c(pooled_hallmarks, halls)
  }  
}

# 2. Plot fig: How many hallmarks is a feature associated with?
tiff(
  paste(
    cancer,
    'hallmark_fts', 
    unlist(strsplit(platform_of_interest, ':'))[2],
    '.tiff',
    sep=''
  ),
  width = 1400,
  height = 1200,
  res = 200,
  compression = "none"
)
df2 <- table(n_hallmarks) %>% as.data.frame()
colnames(df2)<- c('nHallmarks', 'Freq')
ggplot(data=df2, aes(x=nHallmarks, y=Freq, fill='blue')) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=Freq), vjust=-0.3, size=3.5) +
  ggtitle('Histogram - How many hallmarks is a feature associated with?') +
  theme(legend.position = "none")
dev.off()

# 3. Plot fig: What hallmarks are fts most often associated with?
tiff(
  paste(
    cancer,
    'hallmark_hist', 
    unlist(strsplit(platform_of_interest, ':'))[2],
    '.tiff',
    sep=''
  ),
  width = 1400,
  height = 1200,
  res = 200,
  compression = "none"
)
df2 <- sort(table(pooled_hallmarks), decreasing=T) %>% as.data.frame()
colnames(df2)<- c('Hallmarks', 'Freq')
ggplot(data=df2, aes(x=Hallmarks, y=Freq, fill='blue')) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=Freq), hjust=-0.5, size=1.5) +
  ggtitle('What hallmarks are fts most often associated with') +
  theme(text = element_text(size=6)) +
  theme(legend.position = "none") +
  coord_flip()
dev.off()
```


### Section 3: WIP Heatmap with Hallmarks and Feature Importance (Top 5)


```{r}
# Get top data
hallmark_data <- sort(table(pooled_hallmarks), decreasing=T)

# Find top hallmarks
print('Heatmap below will use these hallmarks (order matches with heatmap)')
top <- hallmark_data[1:5]
top <- names(top)
```

**NEED TO MANUALLY UPDATE THIS CODE CHUNK**

```{r}
print('need to manually update this')


gene_set_size <- function(input_hallmark){
  # subset for hallmark rows
  tab<- mappings[mappings['gs_name']==input_hallmark,]
  # count unique gene symbols
  genes <- unique(unlist(tab['human_gene_symbol']))
  return(length(genes))
}

for (h in top){
  print(h)
  print(gene_set_size(h))
}

```

```{r}
# TODO add these as argument input into function
header_jadbio <- models['JADBIO']
header_cforest <- models['CForest']
header_aklimate <- models['AKLIMATE']
header_subscope <- models['SubSCOPE']
header_skgrid <- models['SKGrid']

subtype_ha <- subtype_annotation
```


```{r}
# Function for pulling feature importance scores
normalize_data <- function(df, team_df){
  ## Purpose: min-max normalization and order features to match heatmap
  ## Input df that contains columns c("features","importance")
  ## Input df of all teams
  ## Return vector of normalized values

  # Min max normalize importance scores
  subset_df <- df[,c('features', 'importance')]
  minmax_vals <- mm_normal(subset_df$importance)
  subset_df <- cbind(subset_df, minmax_vals)
  

  # Map and order ft importances to pooled ft order
  minmax_norm <- c()
  test_ft_order <- c()
  all_features <- subset_df$features
  for (pooled_ft in team_df$featureID){
    if (pooled_ft %in% all_features){
      mm <- subset_df[subset_df$'features'==pooled_ft,]$minmax_vals
      minmax_norm <- c(minmax_norm, mm)
      test_ft_order <- c(test_ft_order, pooled_ft)
    } else {
      minmax_norm <- c(minmax_norm, 0)
      test_ft_order <- c(test_ft_order, pooled_ft)
    }
  }
  return(minmax_norm)
}
```




```{r}
setwd(outdir)

# TODO make less janky so autopopulates the annotation names 
if (cancer == 'LGGGBM' & top == c("INTERFERON_GAMMA_RESPONSE", "MTORC1_SIGNALING", "IL2_STAT5_SIGNALING", "INFLAMMATORY_RESPONSE", "ESTROGEN_RESPONSE_LATE") ){
  vals_1 <- build_hallmark_vect(top[1],ftnames_order, platform_of_interest)
  vals_2 <- build_hallmark_vect(top[2],ftnames_order, platform_of_interest)
  vals_3 <- build_hallmark_vect(top[3],ftnames_order, platform_of_interest)
  vals_4 <- build_hallmark_vect(top[4],ftnames_order, platform_of_interest)
  vals_5 <- build_hallmark_vect(top[5],ftnames_order, platform_of_interest)
  
  # Build annotation bars of teams feature sets. 
  # 1A. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  
  
  ########
  # 1B. Create MinMax Values
  ########
  aklimate_minmax <- normalize_data(imp_aklimate, team_df)
  jadbio_minmax <- normalize_data(imp_jd, team_df)


  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  
  # Build annotation
  col_annot <- HeatmapAnnotation(
    # A. ft binary membership
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    CloudForest = cforest,
    JADBio = jadbio,
    SciKitGrid = skgrid,

    # B. ft importance
    AKLIMATE_importance = aklimate_minmax,
    JADBio_importance_v7 = jadbio_minmax,
    
    # C. N teams selected
    PercentSelected= anno_lines(
      mapply('/',team_df$Total, 5) *100,
      height = unit(10, 'mm')
    ),
    
    # TODO need to manually update the name
    # D. Hallmarks
    'InterferonGammaResponse_200' = vals_1,
    'MTORC1Signaling_200' = vals_2,
    'IL2STAT5Signaling_199' = vals_3,
    'InflammatoryResponse_200' = vals_4,
    'EstrogenResponseLate_200' = vals_5, 
    
    col = list(
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      CloudForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      JADBio = c('0' = "#333333", '1' = "#D55B5B"),
      SciKitGrid =  c('0' = "#333333", '1' = "#EFA9A9"),
      AKLIMATE_importance = colorRamp2(c(0, 1), c("#333333", "cadetblue1")),
      JADBio_importance_v7 = colorRamp2(c(0, 1), c("#333333", "#D55B5B")),
      'InterferonGammaResponse_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'MTORC1Signaling_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'IL2STAT5Signaling_199' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'InflammatoryResponse_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'EstrogenResponseLate_200' = c('0' = "#333333", '1' = "darkgoldenrod3")
    ),
    show_legend = FALSE,
    gap = unit(c(0,0,0,0,1,0,1,1,0,0,0,0), 'mm')
  )
  
  # Plot
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    # width = unit(10, 'cm'),
    # height = unit(10, 'cm'),
    name = unlist(strsplit(platform_of_interest, ':'))[2],
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = paste('Features (n=', n_fts, ')', sep=''),
    column_title_gp = gpar(fontsize = 12),
    row_title = paste('Samples (n=', nrow(mat2), ')', sep=''),
    row_title_gp = gpar(fontsize = 12),
    left_annotation = subtype_ha,
    bottom_annotation = col_annot,
    use_raster = TRUE,
    na_col = 'white',
    col = colorRamp2(c(-2, 0, 2), c('blue', 'white', 'red'))
  )
  # Set up saving fig packet
  tiff(
    paste(
      cancer,
      'ht_ftimportance_top5', 
      unlist(strsplit(platform_of_interest, ':'))[2],
      '.tiff',
      sep=''
    ),
    width = 1600,
    height = 1400,
    res = 200,
    compression = "none"
  )
  draw(fig,heatmap_legend_side = c('left'))
  dev.off()

} else if (cancer == 'BRCA' & top == c("ESTROGEN_RESPONSE_EARLY", "ESTROGEN_RESPONSE_LATE", "E2F_TARGETS", "G2M_CHECKPOINT", "MTORC1_SIGNALING" ) ){
  vals_1 <- build_hallmark_vect(top[1],ftnames_order, platform_of_interest)
  vals_2 <- build_hallmark_vect(top[2],ftnames_order, platform_of_interest)
  vals_3 <- build_hallmark_vect(top[3],ftnames_order, platform_of_interest)
  vals_4 <- build_hallmark_vect(top[4],ftnames_order, platform_of_interest)
  vals_5 <- build_hallmark_vect(top[5],ftnames_order, platform_of_interest)
  
  # Build annotation bars of teams feature sets. 
  # 1A. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  
  ########
  # 1B. Create MinMax Values - for AKLIMATE only
  ########
  aklimate_minmax <- normalize_data(imp_aklimate, team_df)
  jadbio_minmax <- normalize_data(imp_jd, team_df)


  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  
  # Build annotation
  col_annot <- HeatmapAnnotation(
    # A. ft binary membership
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    CloudForest = cforest,
    JADBio = jadbio,
    SciKitGrid = skgrid,

    # B. ft importance
    AKLIMATE_importance = aklimate_minmax,
    JADBio_importance_v7 = jadbio_minmax,
        
    # C. N teams selected
    PercentSelected= anno_lines(
      mapply('/',team_df$Total, 5) *100,
      height = unit(10, 'mm')
    ),
    
    # TODO need to manually update the name + gene set size of each hallmark name + color pallate
    # D. Hallmarks
    'EstrogrenResponseEarly_200' = vals_1,
    'EstrogrenResponseLate_200' = vals_2,
    'E2FTargets_200' = vals_3,
    'G2MCheckpoint_200' = vals_4,
    'MTORC1Signaling_200' = vals_5,  
    
    col = list(
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      CloudForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      JADBio = c('0' = "#333333", '1' = "#D55B5B"),
      SciKitGrid =  c('0' = "#333333", '1' = "#EFA9A9"),
      AKLIMATE_importance = colorRamp2(c(0, 1), c("#333333", "cadetblue1")),
      JADBio_importance_v7 = colorRamp2(c(0, 1), c("#333333", "#D55B5B")),
      'EstrogrenResponseEarly_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'EstrogrenResponseLate_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'E2FTargets_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'G2MCheckpoint_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'MTORC1Signaling_200' = c('0' = "#333333", '1' = "darkgoldenrod3")
    ),
    show_legend = FALSE,
    gap = unit(c(0,0,0,0,1,0,1,1,0,0,0,0), 'mm')
  )
  
  # Plot
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    # width = unit(10, 'cm'),
    # height = unit(10, 'cm'),
    name = unlist(strsplit(platform_of_interest, ':'))[2],
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = paste('Features (n=', n_fts, ')', sep=''),
    column_title_gp = gpar(fontsize = 12),
    row_title = paste('Samples (n=', nrow(mat2), ')', sep=''),
    row_title_gp = gpar(fontsize = 12),
    left_annotation = subtype_ha,
    bottom_annotation = col_annot,
    use_raster = TRUE,
    na_col = 'white',
    col = colorRamp2(c(-2, 0, 2), c('blue', 'white', 'red'))
  )
  # Set up saving fig packet
  tiff(
    paste(
      cancer,
      'ht_ftimportance_top5', 
      unlist(strsplit(platform_of_interest, ':'))[2],
      '.tiff',
      sep=''
    ),
    width = 1600,
    height = 1400,
    res = 200,
    compression = "none"
  )
  draw(fig,heatmap_legend_side = c('left'))
  dev.off()
} else if (cancer == 'GEA' & top == c("IL2_STAT5_SIGNALING", "APICAL_JUNCTION", "EPITHELIAL_MESENCHYMAL_TRANSITION", "HYPOXIA", "MYOGENESIS") ){
  vals_1 <- build_hallmark_vect(top[1],ftnames_order, platform_of_interest)
  vals_2 <- build_hallmark_vect(top[2],ftnames_order, platform_of_interest)
  vals_3 <- build_hallmark_vect(top[3],ftnames_order, platform_of_interest)
  vals_4 <- build_hallmark_vect(top[4],ftnames_order, platform_of_interest)
  vals_5 <- build_hallmark_vect(top[5],ftnames_order, platform_of_interest)
  
  # Build annotation bars of teams feature sets. 
  # 1A. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  
  ########
  # 1B. Create MinMax Values - for AKLIMATE only
  ########
  aklimate_minmax <- normalize_data(imp_aklimate, team_df)
  jadbio_minmax <- normalize_data(imp_jd, team_df)


  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  
  # Build annotation
  col_annot <- HeatmapAnnotation(
    # A. ft binary membership
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    CloudForest = cforest,
    JADBio = jadbio,
    SciKitGrid = skgrid,

    # B. Ft importance
    AKLIMATE_importance = aklimate_minmax,
    JADBio_importance_v7 = jadbio_minmax,
        
    # C. N teams selected
    PercentSelected= anno_lines(
      mapply('/',team_df$Total, 5) *100,
      height = unit(10, 'mm')
    ),
    
    # D. Hallmarks
    # TODO need to manually update the name + gene set size of each hallmark name + color pallate
    'IL2STAT5Signaling_199' = vals_1,
    'ApicalJunction_200' = vals_2,
    'EpithelialMesenchymalTransition_200' = vals_3,
    'Hypoxia_200' = vals_4,
    'Myogenesis_200' = vals_5,  
    
    col = list(
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      CloudForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      JADBio = c('0' = "#333333", '1' = "#D55B5B"),
      SciKitGrid =  c('0' = "#333333", '1' = "#EFA9A9"),
      AKLIMATE_importance = colorRamp2(c(0, 1), c("#333333", "cadetblue1")),
      JADBio_importance_v7 = colorRamp2(c(0, 1), c("#333333", "#D55B5B")),
      'IL2STAT5Signaling_199' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'ApicalJunction_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'EpithelialMesenchymalTransition_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'Hypoxia_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      'Myogenesis_200' = c('0' = "#333333", '1' = "darkgoldenrod3")
    ),
    show_legend = FALSE,
    gap = unit(c(0,0,0,0,1,0,1,1,0,0,0,0), 'mm')
  )
  
  # Plot
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    # width = unit(10, 'cm'),
    # height = unit(10, 'cm'),
    name = unlist(strsplit(platform_of_interest, ':'))[2],
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = paste('Features (n=', n_fts, ')', sep=''),
    column_title_gp = gpar(fontsize = 12),
    row_title = paste('Samples (n=', nrow(mat2), ')', sep=''),
    row_title_gp = gpar(fontsize = 12),
    left_annotation = subtype_ha,
    bottom_annotation = col_annot,
    use_raster = TRUE,
    na_col = 'white',
    col = colorRamp2(c(-2, 0, 2), c('blue', 'white', 'red'))
  )
  # Set up saving fig packet
  tiff(
    paste(
      cancer,
      'ht_ftimportance_top5', 
      unlist(strsplit(platform_of_interest, ':'))[2],
      '.tiff',
      sep=''
    ),
    width = 1600,
    height = 1400,
    res = 200,
    compression = "none"
  )
  draw(fig,heatmap_legend_side = c('left'))
  dev.off()
  
}
```