---
title: "Molecular Heatmap - Exact Feature Match"
author: "Jordan Lee"
date: "1/12/2021"
output: html_document
---

# TMP GDAN working group 
## Figure 4 Suggestion

### Question

We see (in other figures) that different teams have similar high classification performance. Is this similar performance due to selection of the same feature sets?

### Rationale

If we see teams picking different features, this suggests that there are multiple biomarkers that can predict cancer subtype well. An alternative explanation if we see teams picking different features, is that perhaps there is some relationship (correlation, biological pathway connection, etc.) between features so in a sense the teams are picking the "same" features if they are correlated.

In this analysis we will address: are groups picking the same features? What is the relationship?


### Analysis

The preprocessing and generation of these files was conducted in a custom python pipeline. A summary of the pipeline is:

1. Pick a cancer cohort (here we picked breast cancer)
2. Select the best model from each of the 5 teams (JadBIO, SubSCOPE, AKLIMATE, Sci-Kit Grid, and Cloud Forest). Based on highest mean overall weighted F1 score
3. Pull the corresponding feature set from each model
4. Map features to molecular file (based on Theo's tarball)
5. Cluster feature columns, no clustering of sample rows Clustering done on molecular file

Note in calculating z-scores, NAs were omitted

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(data.table))
# install.packages('testit')
suppressPackageStartupMessages(library(testit))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
```

```{r}
###### Hardcoded 
cancer <- 'BRCA' # 'GEA', 'LGGGBM', 'BRCA'

outdir ='../data/figure_panel_b/'
df <- fread(paste('../data/figure_panel_b/', cancer, '_fts_by_VALUE.tsv', sep='')) %>% as.data.frame()
df_fts <- fread(paste('../data/figure_panel_b/', cancer, '_fts_by_TEAM.tsv', sep=''))%>% as.data.frame()
mappings <- fread('../data/figure_panel_b/hallmarks.tsv') %>% as.data.frame()
######
```

### Section 1: Heatmaps for each data type

Scaling data - METH, GEXP, and MIR

Non-scaled data - MUTA, CNVR

I want to see the differences between features. Data was scaled - all samples for a given feature have mean 0 and sd 1. then ran through normal pipeline

<details><summary>Functions</summary><p>

```{r}
model2team <- function(df){
  # Read team model names and map to team name
  headers_vector <- c()
  names_vector <- c()
  
  models <- colnames(df)[!colnames(df) %in% c('featureID', 'Total')]
  
  for (header in models){
    if (grepl('gnosis', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'JADBIO')
    } 
    else if (grepl('CF', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'CForest')
    }
    else if (grepl('nn_jg', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'SubSCOPE')
    }
    else if (grepl('AKLIMATE', header, fixed=TRUE)){
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'AKLIMATE')
    }
    else {
      headers_vector <- c(headers_vector, header)
      names_vector <- c(names_vector, 'SKGrid')
    }  
  }
  names(headers_vector) <- names_vector
  return(headers_vector)
}


get_colors <- function(df){
  # Pull color codes based on total number of subtypes
  nsubs <- df %>% 
    select(Labels) %>% 
    unique() %>%
    nrow() %>% 
    as.integer()
  if (nsubs == 4){
    color_codes <- c(
      "1" = 'salmon4',
      "2"='red3',
      "3"='orangered',
      "4"='orange1'
    )
  } else if (nsubs == 7){
    color_codes <- c(
      "1" = '#F58748',
      "2" ='#A95757',
      "3"='#4A2918',
      "4"='red3',
      "5"='orangered',
      "6"= 'salmon',
      "7" = 'orange1'
    )  
  }
  return(color_codes)
}



get_base_heatmap <- function(prefix, cancer, header_jadbio, header_cforest, header_aklimate, header_subscope, header_skgrid){
  # A. Order by subtype
  df_transform <- df %>% arrange(Labels)
  
  # B. Column annotation
  subtypestring <- paste(cancer, '_', sep='')
  
  subtype_ha <- rowAnnotation(
    Subtype = 
      select(df_transform, Labels) %>%
      mutate_all(funs(str_replace(., subtypestring, ""))) %>%
      as.matrix(), 
    na_col = 'grey', 
    col = list(
      Subtype = get_colors(df)
    ),
    show_annotation_name = FALSE,
    simple_anno_size = unit(2, "mm")
  )
  
  # C. Select data type
  df_transform <- df_transform %>%
      select(-Labels) %>%
      select(-all_of(cancer)) %>%
      select(starts_with(prefix))
  mat <- df_transform %>%
      as.matrix() %>%
      t()
  print(prefix)
  n_fts <- nrow(mat)
  print(n_fts)
  
  # 1. Heatmap that applies clustering we want
  #    second heatmap same row/col order but with added annotation bars
  
  # scale if appropriate 
  if (prefix %in% yes_scale){
    mat <- scale(t(mat), center=TRUE, scale=TRUE) #z-scores == each ft row will have mean 0, sd 1. omit NAs
  } else {
    mat <- t(mat) #flip for heatmap looks
  }
  
  fig <- Heatmap(
    mat,
    name = 'first heatmap',
    cluster_rows = FALSE,
    clustering_distance_columns = "euclidean",
    clustering_method_columns = "ward.D",
    cluster_columns = TRUE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = 'Samples',
    row_title = 'Features',
    right_annotation = subtype_ha
  )
  # draw(fig) # dont need to draw
  
  
  #### 
  # Add team annotation bar
  ####
  # Ordering
  # 1. Get order of features post heatmap clustering
  heatmap_order <- column_order(fig) # index vector
  ftnames_order <- c() # featurename vector
  for (i in heatmap_order){
    add_ft <- colnames(mat)[i]
    ftnames_order <- c(ftnames_order, add_ft)
  }
  # print(ftnames_order)
  # 2. Get new matrix that is ordered by heatmap clustering
  mat2 <- mat[,match(ftnames_order, colnames(mat))]
  
  
  
  # Build annotation bars of teams feature sets. 
  # 1. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  team_list <- HeatmapAnnotation(
    JADBIO = jadbio,
    CForest = cforest,
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    SKGrid = skgrid,
    col = list(
      JADBIO = c('0' = "#333333", '1' = "palevioletred1"),
      CForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      SKGrid =  c('0' = "#333333", '1' = "rosybrown2")
    ),
    show_legend = FALSE,
    nTeamsSelected= anno_barplot(team_df$Total, bar_width = 1)
  )
  
  # 3. Heatmap
  # Set up saving fig packet
  pdf(paste(cancer, 'ht_base', unlist(strsplit(prefix, ':'))[2],'.pdf',sep='_'), onefile=FALSE)
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    name = prefix,
    cluster_rows = FALSE,
    clustering_distance_columns = "euclidean",
    clustering_method_columns = "ward.D",
    column_order = ftnames_order,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = 'Samples',
    row_title = 'Features',
    right_annotation = subtype_ha,
    top_annotation = team_list,
  )
  draw(fig)
  dev.off()
  
  
  #####
  # unit tests
  #####
  # sanity check 1 - that the row orders are the same in the two heatmaps
  assert('Assertion Error: order of feature rows on heatmap and final heatmap not same', colnames(mat2) == ftnames_order)
  # sanity check 2 - directly compare heatmap rows
  heatmap_order <- column_order(fig) # index vector
  ftnames_order2 <- c() # featurename vector
  for (i in heatmap_order){
    add_ft <- colnames(mat2)[i]
    ftnames_order2 <- c(ftnames_order2, add_ft)
  }
  assert('Assertion Error: order of feature rows on heatmap and final heatmap not same', ftnames_order2 == ftnames_order)
  
  print(paste('Distance metric = ', fig@row_dend_param$distance, '. Method = ', fig@row_dend_param$method, sep=' '))
  
  # Assign to 'output' variables
  return(list('results_matrix' = mat2, 'results_ft_order' = ftnames_order, 'subtype_annotation' = subtype_ha))
}
```





</details>

<details><summary>Code and figures for all data types</summary><p>

```{r mixed scaled z score loop for all data types}
####
# Create Heatmap for all data types
####
# Define platform for hallmark heatmap
if (cancer == 'BRCA'){
  platform_of_interest <- 'N:GEXP'
} else if (cancer == 'LGGGBM'){
  platform_of_interest <- 'N:METH'
} else if (cancer == 'GEA'){
  platform_of_interest <- 'N:METH'
}


# Build list of data types present
if (cancer == 'BRCA' | cancer == 'GEA' ){
  platforms <- c('N:METH', 'N:MIR', 'I:CNVR', 'B:MUTA', 'N:GEXP')
} else if (cancer == 'LGGGBM'){
  platforms <- c('N:METH', 'I:CNVR', 'B:MUTA', 'N:GEXP')
}



# Set which to scale
yes_scale <- c('N:METH', 'N:GEXP','N:MIR') # which fts to scale

# get models
models <- model2team(df_fts)

# Set up saving fig packet
setwd(outdir)
for (prefix in platforms){
  # Create Base Heatmap
  results_list <- get_base_heatmap(
    prefix, 
    cancer, 
    models['JADBIO'],
    models['CForest'],
    models['AKLIMATE'],
    models['SubSCOPE'],
    models['SKGrid']
  )
  
  # If platform for Hallmarks Heatmap then save global variables - input for Section 2
  if (prefix == platform_of_interest){
    print('want to use this one')
    print(prefix)
    mat2 <- results_list[['results_matrix']]
    ftnames_order <- results_list[['results_ft_order']]  
    subtype_annotation <- results_list[['subtype_annotation']]
  }
}


```



</details>


### Section 2: Explore the hallmarks associated with GEXP features

One annotation bar (top hallmarks only, all others are not colored)


<details><summary>2. Explore Hallmarks</summary><p>

```{r explore hallmarks}
######
# Prep
# look up if in hallmark mappings table
n_hallmarks <- c() # exploratory
pooled_hallmarks <- c() # exploratory

if (platform_of_interest == 'N:GEXP'){
  for (feature in ftnames_order){
    # 1. Preprocess - to gene symbol
    GENE <- unlist(strsplit(feature, '::'))[2]
    GENE <- unlist(strsplit(GENE, ':'))[1]
    # 2. Hallmark Mapping
    halls <- mappings[mappings$human_gene_symbol==GENE,]$gs_name %>% as.vector()
    # Multiple Hallmarks
    n_hallmarks <- c(n_hallmarks, length(halls))
    pooled_hallmarks <- c(pooled_hallmarks, halls)
  }  
} else if (platform_of_interest == 'N:METH'){
  for (feature in ftnames_order){
    # 1. Preprocess - to gene symbol
    GENE <- unlist(strsplit(feature, ':'))[4]
    # 2. Hallmark Mapping
    halls <- mappings[mappings$human_gene_symbol==GENE,]$gs_name %>% as.vector()
    # Multiple Hallmarks
    n_hallmarks <- c(n_hallmarks, length(halls))
    pooled_hallmarks <- c(pooled_hallmarks, halls)
  }  
}



# How many hallmarks is a feature associated with?
# Set up saving fig packet
setwd(outdir)
pdf(paste(cancer, 'hallmark_fts', unlist(strsplit(platform_of_interest, ':'))[2],'.pdf',sep='_'), onefile=FALSE)

df2 <- table(n_hallmarks) %>% as.data.frame()
colnames(df2)<- c('nHallmarks', 'Freq')
ggplot(data=df2, aes(x=nHallmarks, y=Freq, fill='blue')) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=Freq), vjust=-0.3, size=3.5) +
  ggtitle('Histogram - How many hallmarks is a feature associated with?') +
  theme(legend.position = "none")
dev.off()

# What hallmarks are fts most often associated with?
# Set up saving fig packet
pdf(paste(cancer, 'hallmark_hist', unlist(strsplit(platform_of_interest, ':'))[2],'.pdf',sep='_'), onefile=FALSE)

df2 <- sort(table(pooled_hallmarks), decreasing=T) %>% as.data.frame()
colnames(df2)<- c('Hallmarks', 'Freq')
ggplot(data=df2, aes(x=Hallmarks, y=Freq, fill='blue')) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=Freq), hjust=-0.5, size=1.5) +
  ggtitle('What hallmarks are fts most often associated with') +
  theme(text = element_text(size=6)) +
  theme(legend.position = "none") +
  coord_flip()
dev.off()
```

</details>

<details><summary>3. Heatmap with Hallmarks (Top 5)</summary><p>

```{r}
####
# function
####
build_hallmark_vect <- function(hallmark, ftnames_order, platform_of_interest){
  fts_checked <- c() # sanity check
  hallmark_present <- c()
  
  for (feature in ftnames_order){
    # 1. Preprocess - to gene symbol
    if (platform_of_interest == 'N:GEXP'){
      GENE <- unlist(strsplit(feature, '::'))[2]
      GENE <- unlist(strsplit(GENE, ':'))[1]
    } else if (platform_of_interest == 'N:METH'){
      GENE <- unlist(strsplit(feature, ':'))[4]
    }

    # 2. Hallmark Mapping
    halls <- mappings[mappings$human_gene_symbol==GENE,]$gs_name %>% as.vector()
    # If hallmark present
    if (hallmark %in% halls){
      i <- match(hallmark, halls)
      hallmark_present <- c(hallmark_present, 1)
      fts_checked <- c(fts_checked, feature)
    }
    else {
      hallmark_present <- c(hallmark_present, 0)
      fts_checked <- c(fts_checked, feature)
    }
  }
  return(hallmark_present)
}
```

```{r}
# Get top data
hallmark_data <- sort(table(pooled_hallmarks), decreasing=T)

# Find top hallmarks
print('Heatmap below will use these hallmarks (order matches with heatmap)')
top <- hallmark_data[1:5]
top <- names(top)
```

#### Need to manually update this code chunk

```{r}
print('need to manually update this')


gene_set_size <- function(input_hallmark){
  # subset for hallmark rows
  tab<- mappings[mappings['gs_name']==input_hallmark,]
  # count unique gene symbols
  genes <- unique(unlist(tab['human_gene_symbol']))
  return(length(genes))
}

for (h in top){
  print(h)
  print(gene_set_size(h))
}

```

```{r to update}
# TODO add these as argument input into function
header_jadbio <- models['JADBIO']
header_cforest <- models['CForest']
header_aklimate <- models['AKLIMATE']
header_subscope <- models['SubSCOPE']
header_skgrid <- models['SKGrid']

subtype_ha <- subtype_annotation

```

```{r hallmarks}
# TODO make less janky so autopopulates the annotation names 
if (cancer == 'LGGGBM' & top == c("INTERFERON_GAMMA_RESPONSE", "MTORC1_SIGNALING", "IL2_STAT5_SIGNALING", "INFLAMMATORY_RESPONSE", "ESTROGEN_RESPONSE_LATE") ){
  vals_1 <- build_hallmark_vect(top[1],ftnames_order, platform_of_interest)
  vals_2 <- build_hallmark_vect(top[2],ftnames_order, platform_of_interest)
  vals_3 <- build_hallmark_vect(top[3],ftnames_order, platform_of_interest)
  vals_4 <- build_hallmark_vect(top[4],ftnames_order, platform_of_interest)
  vals_5 <- build_hallmark_vect(top[5],ftnames_order, platform_of_interest)
  
  # Build annotation bars of teams feature sets. 
  # 1. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  
  # Build annotation
  col_annot <- HeatmapAnnotation(
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    CForest = cforest,
    JADBIO = jadbio,
    SKGrid = skgrid,
    
    nTeamsSelected= anno_barplot(team_df$Total, bar_width = 1, gp = gpar(fill='blue', col='grey38') ),
    
    # TODO need to manually update the name + gene set size of each hallmark name + color pallate
    'InterferonGammaResponse_200' = vals_1,
    MTORC1Signaling_200 = vals_2,
    IL2STAT5Signaling_199 = vals_3,
    InflammatoryResponse_200 = vals_4,
    EstrogenResponseLate_200 = vals_5,  
    
    col = list(
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      CForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      JADBIO = c('0' = "#333333", '1' = "palevioletred1"),
      SKGrid =  c('0' = "#333333", '1' = "rosybrown2"),
      'InterferonGammaResponse_200' = c('0' = "#333333", '1' = "darkgoldenrod3"),
      MTORC1Signaling_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      IL2STAT5Signaling_199 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      InflammatoryResponse_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      EstrogenResponseLate_200 = c('0' = "#333333", '1' = "darkgoldenrod3")
    ),
    show_legend = FALSE,
    gap = unit(c(0,0,0,0,1,1,0,0,0,0), 'mm')
  )
  
  # Plot
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    name = unlist(strsplit(platform_of_interest, ':'))[2],
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = 'Samples',
    row_title = 'Features',
    right_annotation = subtype_ha,
    bottom_annotation = col_annot,
    use_raster = TRUE
  )
  # Set up saving fig packet
  setwd(outdir)
  pdf(paste(cancer, 'ht_top5', unlist(strsplit(platform_of_interest, ':'))[2],'.pdf',sep='_'), onefile=FALSE)
  draw(fig)
  dev.off()
} else if (cancer == 'BRCA' & top == c("ESTROGEN_RESPONSE_EARLY", "ESTROGEN_RESPONSE_LATE", "E2F_TARGETS", "G2M_CHECKPOINT", "MTORC1_SIGNALING" ) ){
  vals_1 <- build_hallmark_vect(top[1],ftnames_order, platform_of_interest)
  vals_2 <- build_hallmark_vect(top[2],ftnames_order, platform_of_interest)
  vals_3 <- build_hallmark_vect(top[3],ftnames_order, platform_of_interest)
  vals_4 <- build_hallmark_vect(top[4],ftnames_order, platform_of_interest)
  vals_5 <- build_hallmark_vect(top[5],ftnames_order, platform_of_interest)
  
  # Build annotation bars of teams feature sets. 
  # 1. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  
  # Build annotation
  col_annot <- HeatmapAnnotation(
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    CForest = cforest,
    JADBIO = jadbio,
    SKGrid = skgrid,
    
    nTeamsSelected= anno_barplot(team_df$Total, bar_width = 1, gp = gpar(fill='blue', col='grey38') ),
    
    # TODO need to manually update the name + gene set size of each hallmark name + color pallate
    EstrogrenResponseEarly_200 = vals_1,
    EstrogrenResponseLate_200 = vals_2,
    E2FTargets_200 = vals_3,
    G2MCheckpoint_200 = vals_4,
    MTORC1Signaling_200 = vals_5,  
    
    col = list(
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      CForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      JADBIO = c('0' = "#333333", '1' = "palevioletred1"),
      SKGrid =  c('0' = "#333333", '1' = "rosybrown2"),
      EstrogrenResponseEarly_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      EstrogrenResponseLate_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      E2FTargets_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      G2MCheckpoint_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      MTORC1Signaling_200 = c('0' = "#333333", '1' = "darkgoldenrod3")
    ),
    show_legend = FALSE,
    gap = unit(c(0,0,0,0,1,1,0,0,0,0), 'mm')
  )
  
  # Plot
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    name = unlist(strsplit(platform_of_interest, ':'))[2],
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = 'Samples',
    row_title = 'Features',
    right_annotation = subtype_ha,
    bottom_annotation = col_annot,
    use_raster = TRUE
  )
  # Set up saving fig packet
  setwd(outdir)
  pdf(paste(cancer, 'ht_top5', unlist(strsplit(platform_of_interest, ':'))[2],'.pdf',sep='_'), onefile=FALSE)
  draw(fig)
  dev.off()
} else if (cancer == 'GEA' & top == c("IL2_STAT5_SIGNALING", "APICAL_JUNCTION", "EPITHELIAL_MESENCHYMAL_TRANSITION", "HYPOXIA", "MYOGENESIS") ){
  vals_1 <- build_hallmark_vect(top[1],ftnames_order, platform_of_interest)
  vals_2 <- build_hallmark_vect(top[2],ftnames_order, platform_of_interest)
  vals_3 <- build_hallmark_vect(top[3],ftnames_order, platform_of_interest)
  vals_4 <- build_hallmark_vect(top[4],ftnames_order, platform_of_interest)
  vals_5 <- build_hallmark_vect(top[5],ftnames_order, platform_of_interest)
  
  # Build annotation bars of teams feature sets. 
  # 1. df of all teams. match ft order in heatmap
  team_df<- df_fts %>% filter(featureID %in% ftnames_order) %>% arrange(match(featureID, ftnames_order))
  # 2. Pull just the team of interest
  jadbio <- team_df %>% pull(header_jadbio) %>% as.character()
  cforest <- team_df %>% pull(header_cforest) %>% as.character()
  aklimate <- team_df %>% pull(header_aklimate) %>% as.character()
  subscope <- team_df %>% pull(header_subscope) %>% as.character()
  skgrid <- team_df %>% pull(header_skgrid) %>% as.character()
  
  # Build annotation
  col_annot <- HeatmapAnnotation(
    AKLIMATE = aklimate,
    SubSCOPE = subscope,
    CForest = cforest,
    JADBIO = jadbio,
    SKGrid = skgrid,
    
    nTeamsSelected= anno_barplot(team_df$Total, bar_width = 1, gp = gpar(fill='blue', col='grey38') ),
    
    # TODO need to manually update the name + gene set size of each hallmark name + color pallate
    IL2STAT5Signaling_199 = vals_1,
    ApicalJunction_200 = vals_2,
    EpithelialMesenchymalTransition_200 = vals_3,
    Hypoxia_200 = vals_4,
    Myogenesis_200 = vals_5,  
    
    col = list(
      AKLIMATE =  c('0' = "#333333", '1' = "cadetblue1"),
      SubSCOPE =  c('0' = "#333333", '1' = "palegreen2"),
      CForest =  c('0' = "#333333", '1' = "mediumpurple1"),
      JADBIO = c('0' = "#333333", '1' = "palevioletred1"),
      SKGrid =  c('0' = "#333333", '1' = "rosybrown2"),
      IL2STAT5Signaling_199 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      ApicalJunction_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      EpithelialMesenchymalTransition_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      Hypoxia_200 = c('0' = "#333333", '1' = "darkgoldenrod3"),
      Myogenesis_200 = c('0' = "#333333", '1' = "darkgoldenrod3")
    ),
    show_legend = FALSE,
    gap = unit(c(0,0,0,0,1,1,0,0,0,0), 'mm')
  )
  
  # Plot
  fig <- Heatmap(
    mat2, #each col will have mean 0, sd 1
    name = unlist(strsplit(platform_of_interest, ':'))[2],
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    column_title = 'Samples',
    row_title = 'Features',
    right_annotation = subtype_ha,
    bottom_annotation = col_annot,
    use_raster = TRUE
  )
  # Set up saving fig packet
  setwd(outdir)
  pdf(paste(cancer, 'ht_top5', unlist(strsplit(platform_of_interest, ':'))[2],'.pdf',sep='_'), onefile=FALSE)
  draw(fig)
  dev.off()
  
}
```

</details>